Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2019-11-01 manuscript/climate_sensitivity_draft.Rmd
Update climate_sensitivity_draft.Rmd

Our results reveal how tree size, microhabitat, and hydraulic traits shaped tree growth responses across three droughts in a temperate deciduous forest (Table 1). The tendencey for larger trees to suffer more, observed here as in forests around the world (**REFS**), was driven primarily by their height. There was a marginal additional effect of crown exposure, with the most exposed and the most suppressed trees suffering most--consistent with observations of both greater drought sensitivity of exposed trees (e.g., [@suarez_factors_2004]; [@scharnweber_confessions_2019]; MORE??) and greater sensitivity of suppressed and crowded individuals (**REFS- **). There was no evidence that root water access increased drougth resistance; in contrast, trees in wetter topographic positions suffered more (consistent with [@zuleta_drought-induced_2017]), and larger rooting volume provided no advantage in the drier microenvironments. The lower drought resistance of larger trees was not driven by any tendency for the canopy to be dominated by more drought-sensitive species. Drought-sensitive species were not linked to $LMA$ and $WD$ but were predicted by hydraulic traits ($PLA_{dry}$, $\pi_{tlp}$, and $XP$), which is physiologically logical (**REFS**[@kannenberg_linking_2019]) but scientifically novel in that $PLA_{dry}$ and $\pi_{tlp}$ have not previously been linked to drought growth responses. The direction of these responses was mostly consistent across droughts, indicating that they were driven by fundamental physiological mechanisms; however, the strengths of each predictor varied across droughts, indicating that specific drought characteristics interact with tree size, microenvironment, and traits to shape which individuals suffer most. These findings significantly advance our knowledge as to the factors that confer vulnerability or resilience on trees during drought.

>>>>

Our results reveal how tree size, microhabitat, and hydraulic traits shaped tree growth responses across three droughts in a temperate deciduous forest (Table 1). The tendencey for larger trees to suffer more, observed here as in forests around the world [@bennett_larger_2015], was driven primarily by their height. There was a marginal additional effect of crown exposure, with the most exposed and the most suppressed trees suffering most--consistent with observations of both greater drought sensitivity of exposed trees (e.g., [@suarez_factors_2004]; [@scharnweber_confessions_2019]) and greater sensitivity of suppressed and crowded individuals (**REFS**). There was no evidence that root water access increased drougth resistance; in contrast, trees in wetter topographic positions suffered more (consistent with [@zuleta_drought-induced_2017]), and larger rooting volume provided no advantage in the drier microenvironments. The lower drought resistance of larger trees was not driven by any tendency for the canopy to be dominated by more drought-sensitive species. Drought-sensitive species were not linked to $LMA$ and $WD$ but were predicted--at least some of the time--by hydraulic traits ($PLA_{dry}$, $\pi_{tlp}$, and $XP$), which is physiologically logical (**REFS**[@kannenberg_linking_2019]) but scientifically novel in that $PLA_{dry}$ and $\pi_{tlp}$ have not previously been linked to drought growth responses. The direction of these responses was mostly consistent across droughts, indicating that they were driven by fundamental physiological mechanisms; however, the strengths of each predictor varied across droughts, indicating that specific drought characteristics interact with tree size, microenvironment, and traits to shape which individuals suffer most. These findings significantly advance our knowledge as to the factors that confer vulnerability or resilience on trees during drought.
